---
title: "US Electricity"
output: rmarkdown::html_vignette
vignette: >
  %\VignetteIndexEntry{US Electricity}
  %\VignetteEngine{knitr::rmarkdown}
  %\VignetteEncoding{UTF-8}
---

```{r, include = FALSE}
knitr::opts_chunk$set(
  fig.width=8, 
  fig.height=5,
  collapse = TRUE,
  warning = FALSE, 
  message = FALSE,
  comment = "#>"
)
```

The USgrid R package provides a set of high frequency (hourly) time-series datasets, describing the demand and generation of electricity in the US. That includes the following series:

* `US_elec`  - the total hourly demand and supply (generation) for electricity in the US since July 2015

* `US_source` - the hourly demand for electricity in the US by energy source (natural gas, coal, solar, etc.) since July 2018

* `Cal_elec` - The California subregion hourly demand by operator since July 2018

## Data structure

All the series are in [tsibble](https://tsibble.tidyverts.org/index.html) format

## The US total demand and supply for electricity

The `US_elec` series describes the total hourly demand and supply (generation) of electricity in the US since July 2015:

```{r }
library(USgrid)

data("US_elec")

head(US_elec)

```

Where the `type` variable defines the series key (i.e., demand or generation). As you can see in the plot below, the generation of electricity is relatively close to the demand:


```{r}
library(plotly)

plot_ly(data = US_elec,
        x = ~ date_time,
        y = ~ series,
        color = ~ type,
        line = list(width = 1),
        type = "scatter",
        mode = "lines") %>%
  layout(title = "US Electricity Generation by Source",
         yaxis = list(title = "Mwh"),
         xaxis = list(title = "Source: US Energy Information Administration (Nov 2019)"))
```


## Total demand by energy source

The `US_source` provides a break down of the hourly generation by energy source
